---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# streets

<!-- badges: start -->
<!-- badges: end -->

The goal of streets is to provide functions for working with paths through 2D grids.

## Installation

You can install the development version of streets from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mncube/streets")
```

## Load streets

```{r}
library(streets)
```

## Manhattan Paths

On a 2D rectangular grid, all paths from the origin (0,0) to the destination 
(i,j) have minimum Manhattan Distance if only upward and rightward steps are 
allowed.  The number of unique paths meeting this condition (Manhattan Paths) can
be computed in R as choose(i + j, j) or as factorial(i+j)/(factorial(i)*factorial(j)).
Two example paths with destination (4,4) are plotted below.

```{r ManhattanPathsPlot}
# Define the grid and paths
path1 <- rbind(c(0, 0), c(1, 0), c(1, 1), c(2, 1), c(3, 1), c(4, 1), c(4, 2), c(4, 3), c(4, 4))
path2 <- rbind(c(0, 0), c(0, 1), c(0, 2), c(0, 3), c(1, 3), c(2, 3), c(3, 3), c(3, 4), c(4, 4))

# Set up plot parameters
plot(NULL, xlim = c(0, 4), ylim = c(0, 4), xlab = "i", ylab = "j", xaxs = "i", yaxs = "i", asp = 1)
abline(h = 0:4, v = 0:4, col = "lightgray", lty = 3)

# Plot the paths
lines(path1, col = "blue", lwd = 2)
lines(path2, col = "red", lwd = 2)
```


The mpaths function computes the number of unique Manhattan Paths from the origin
to the destination.

```{r}
mpaths(destination = c(4,4))
```


## Target Paths

Target Paths are similar to Manhattan Paths but they force the paths to pass
through a target location.  In the example below, two target paths on a grid
with destination (4,4) both pass through the target location (2,2).

```{r TargetPathsPlot}
# Define the grid and paths
path1 <- rbind(c(0, 0), c(1, 0), c(2, 0), c(2, 1), c(2, 2), c(3, 2), c(4, 2), c(4, 3), c(4, 4))
path2 <- rbind(c(0, 0), c(0, 1), c(1, 1), c(1, 2), c(2, 2), c(2, 3), c(3, 3), c(3, 4), c(4, 4))

# Set up plot parameters
plot(NULL, xlim = c(0, 4), ylim = c(0, 4), xlab = "i", ylab = "j", xaxs = "i", yaxs = "i", asp = 1)
abline(h = 0:4, v = 0:4, col = "lightgray", lty = 3)

# Plot the paths
lines(path1, col = "blue", lwd = 2)
lines(path2, col = "red", lwd = 2)

# Plot target location
points(2, 2, col = "green", pch = 19, cex = 1)
```


The total number of target paths to a fixed destination can be computed with the
tpaths function.

```{r}
tpaths(destination = c(5,4), target = c(3,2))
```


### Target Paths Probability Functions

The tpaths function and the mpaths function can be used to define probability
distribution function.

The dtpaths function is a probability mass function where the number of target 
paths passing through a target location is the event of interest while the sum of 
target paths across all coordinates on the grid is the sample space. The first
example below shows the symmetry on a square grid.

```{r dtpaths_symmetry}
dtpaths(c(5,5), c(2,3))
dtpaths(c(5,5), c(3,2))
```


The next example shows normalization for dtpaths

```{r dtpaths_normalization}
  probs_tpaths <- 0
  for (x in 0:10){
    for (y in 0:10){
      probs_tpaths <- probs_tpaths + dtpaths(c(10,10), c(x,y))
    }
  }

print(probs_tpaths)
```


The ptpaths function is a cumulative distribution function for the dtpaths 
probability mass function.  The distribution is obtained by iterating through all 
possible target coordinates up to and including (ti, tj), computing dtpaths at 
each of these target coordinates, and summing the results.

```{r ptpaths_ex}
ptpaths(destination = c(8,8),target = c(4,4))
```


The qtpaths function is a quantile function which takes a probability p and a 
destination (i, j) and returns the coordinates for the target location (ti, tj) 
where the cumulative probability is equal to or just greater than p.  If multiple
target locations are tied, then the target location with minimum Manhattan Distance
from the origin wil be selected.  If there are still ties, then the target location
will be selected based on a "col"umn-wise ordering of the coordinates (j then i)
or based on a "row"-wise ordering of the coordinates (i then j).


```{r qtpaths_ex}
qtpaths(destination = c(5,5), p =.5)

# Enter the target coordinate returned by qtpaths into ptpaths to recover p
ptpaths(destination = c(5,5), target = c(2,5))
```


The rtpaths function is a random variable function that generates random target 
coordinates from the distribution defined by the dtpaths probability mass function.

```{r rtpaths_ex}
rtpaths(destination = c(10,10), n = 3)
```


## Detour Paths

Detour Paths are similar to Target Paths but they force the paths NOT to pass
through a detour location.  In the example below, two detour paths on a grid
with destination (4,4) both DO NOT pass through the detour location (2,2).

```{r DetourPathsPlot}
# Define the grid and paths
path1 <- rbind(c(0, 0), c(1, 0), c(2, 0), c(2, 1), c(3, 1), c(3, 2), c(4, 2), c(4, 3), c(4, 4))
path2 <- rbind(c(0, 0), c(0, 1), c(1, 1), c(1, 2), c(1, 3), c(2, 3), c(3, 3), c(3, 4), c(4, 4))

# Set up plot parameters
plot(NULL, xlim = c(0, 4), ylim = c(0, 4), xlab = "i", ylab = "j", xaxs = "i", yaxs = "i", asp = 1)
abline(h = 0:4, v = 0:4, col = "lightgray", lty = 3)

# Plot the paths
lines(path1, col = "blue", lwd = 2)
lines(path2, col = "red", lwd = 2)

# Plot detour location
points(2, 2, col = "green", pch = 19, cex = 1)
```


The total number of detour paths to a fixed destination can be computed with the
dpaths function.

```{r}
dpaths(destination = c(5,4), detour = c(3,2))
```


Note the relation between mpaths, tpaths, and dpaths

```{r}
mpaths(destination = c(5,4)) == tpaths(destination = c(5,4), target = c(3,2)) +
                                   dpaths(destination = c(5,4), detour = c(3,2))
```


### Detour Paths Probability Functions

The streets package has a probability mass function (ddpaths), a cumulative
probability function (pdpaths), a quantile function (qdpaths), and a random
variable functions (rdpaths) for detour paths.  See examples below.

```{r detourpaths_probabilityfuns}
# Symmetry of the probability mass function
ddpaths(c(5,5), c(2,3))
ddpaths(c(5,5), c(3,2))

# Normalization of the probability mass function
  probs_dpaths <- 0
  for (x in 0:10){
    for (y in 0:10){
      probs_dpaths <- probs_dpaths + ddpaths(c(10,10), c(x,y))
    }
  }

print(probs_dpaths)

# Cumulative distribution
pdpaths(destination = c(8,8), detour = c(4,4))

# Quantile function
qdpaths(destination = c(5,5), p =.5)
pdpaths(destination = c(5,5), detour = c(5,2)) #Recover p

# Random detour location
rdpaths(destination = c(10,10), n = 3)
```


## Branch Paths

Branch Paths force the paths to pass to pass through a branch location; moreover,
at a branch location, the path is allowed to step in all three directions exiting
the location.  As such, Branch Paths are similar to no-U-turn intersections or an
out-degree-3 nodes.  In the example below, a branch path approaches a branch location
from below and exits with a step that does not produce the minimum Manhattan 
Distance to the destination.

```{r BranchPathsPlot}
# Define the grid and path
path1 <- rbind(c(0, 0), c(1, 0), c(2, 0), c(2,1), c(2, 2), c(1, 2), c(1, 3), c(2, 3), 
               c(3, 3), c(3, 4), c(4,4))

# Set up plot parameters
plot(NULL, xlim = c(0, 4), ylim = c(0, 4), xlab = "i", ylab = "j", xaxs = "i", yaxs = "i", asp = 1)
abline(h = 0:4, v = 0:4, col = "lightgray", lty = 3)

# Plot the path
lines(path1, col = "blue", lwd = 2)

# Plot branch location
points(2, 2, col = "green", pch = 19, cex = 1)
```


And in the plot below, a path enters a branch location with a rightward step
and exits with a downward step.

```{r BranchPathsPlot2}
# Define the grid and path
path2 <- rbind(c(0, 0), c(0, 1), c(1, 1), c(1, 2), c(2, 2), c(2, 1), c(3, 1), c(3, 2), c(4, 2), c(4,3), c(4,4))

# Set up plot parameters
plot(NULL, xlim = c(0, 4), ylim = c(0, 4), xlab = "i", ylab = "j", xaxs = "i", yaxs = "i", asp = 1)
abline(h = 0:4, v = 0:4, col = "lightgray", lty = 3)

# Plot the path
lines(path2, col = "red", lwd = 2)

# Plot branch location
points(2, 2, col = "green", pch = 19, cex = 1)
```


The total number of branch paths to a fixed destination can be computed with the
bpaths function.

```{r}
bpaths(destination = c(5,4), branch = c(3,2))
```


Note how branch paths behave along the left and bottom boundaries of the grid

```{r}
# Left Boundary (i.e., bi = 0)
bpaths(destination = c(5,5), branch = c(0,3)) == 
  tpaths(destination = c(5,5), target = c(0,3))

# Bottom Boundary (i.e., bj = 0)
bpaths(destination = c(5,5), branch = c(3,0)) == 
  tpaths(destination = c(5,5), target = c(3,0))
```


### Branch Paths Probability Functions

The streets package has a probability mass function (dbpaths), a cumulative
probability function (pbpaths), a quantile function (qbpaths), and a random
variable function (rbpaths) for branch paths.  See examples below.

```{r branchpaths_probabilityfuns}
# Symmetry of the probability mass function
dbpaths(c(5,5), c(2,3))
dbpaths(c(5,5), c(3,2))

# Normalization of the probability mass function
  probs_bpaths <- 0
  for (x in 0:10){
    for (y in 0:10){
      probs_bpaths <- probs_bpaths + dbpaths(c(10,10), c(x,y))
    }
  }

print(probs_bpaths)

# Cumulative distribution
pbpaths(destination = c(8,8), branch = c(4,4))

# Quantile function
qbpaths(destination = c(5,5), p =.5)
pbpaths(destination = c(5,5), branch = c(3,3)) #Recover p

# Random branch locations
rbpaths(destination = c(10,10), n = 3)
```
