---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# streets

<!-- badges: start -->
<!-- badges: end -->

The goal of streets is to provide functions for working with paths through 2D grids.

## Installation

You can install the development version of streets from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mncube/streets")
```

## Load streets

```{r}
library(streets)
```

## Manhattan Paths

On a 2D rectangular grid, all paths from the origin (0,0) to the destination 
(i,j) have minimum Manhattan Distance if only upward and rightward steps are 
allowed.  The number of unique paths meeting this condition (Manhattan Paths) can
be computed in R as choose(i + j, j) or as factorial(i+j)/(factorial(i)*factorial(j)).
Two example paths are plotted below.

```{r ManhattanPathsPlot}
# Define the grid and paths
destination <- c(4, 4)
grid_size <- destination + 1
path1 <- rbind(c(0, 0), c(1, 0), c(1, 1), c(2, 1), c(3, 1), c(4, 1), c(4, 2), c(4, 3), c(4, 4))
path2 <- rbind(c(0, 0), c(0, 1), c(0, 2), c(0, 3), c(1, 3), c(2, 3), c(3, 3), c(3, 4), c(4, 4))

# Set up plot parameters
plot(NULL, xlim = c(0, 4), ylim = c(0, 4), xlab = "X", ylab = "Y", xaxs = "i", yaxs = "i", asp = 1)
abline(h = 0:4, v = 0:4, col = "lightgray", lty = 3)

# Plot the paths
lines(path1, col = "blue", lwd = 2)
lines(path2, col = "red", lwd = 2)
```

The mpaths function computes the number of unique Manhattan Paths from the origin
to the destination.

```{r}
mpaths(destination = c(4,4))
```
