---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# streets

<!-- badges: start -->
<!-- badges: end -->

The goal of streets is to provide functions for working with paths through 2D grids.

## Installation

You can install the development version of streets from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mncube/streets")
```

## Load streets

```{r}
library(streets)
```

## Manhattan Paths

On a 2D rectangular grid, all paths from the origin (0,0) to the destination 
(i,j) have minimum Manhattan Distance if only upward and rightward steps are 
allowed.  The number of unique paths meeting this condition (Manhattan Paths) can
be computed in R as choose(i + j, j) or as factorial(i+j)/(factorial(i)*factorial(j)).
Two example paths with destination (4,4) are plotted below.

```{r ManhattanPathsPlot}
# Define the grid and paths
path1 <- rbind(c(0, 0), c(1, 0), c(1, 1), c(2, 1), c(3, 1), c(4, 1), c(4, 2), c(4, 3), c(4, 4))
path2 <- rbind(c(0, 0), c(0, 1), c(0, 2), c(0, 3), c(1, 3), c(2, 3), c(3, 3), c(3, 4), c(4, 4))

# Set up plot parameters
plot(NULL, xlim = c(0, 4), ylim = c(0, 4), xlab = "X", ylab = "Y", xaxs = "i", yaxs = "i", asp = 1)
abline(h = 0:4, v = 0:4, col = "lightgray", lty = 3)

# Plot the paths
lines(path1, col = "blue", lwd = 2)
lines(path2, col = "red", lwd = 2)
```


The mpaths function computes the number of unique Manhattan Paths from the origin
to the destination.

```{r}
mpaths(destination = c(4,4))
```


## Target Paths

Target Paths are similar to Manhattan Paths but they force the paths to pass
through a target location.  In the example below, two target paths on a grid
with destination (4,4) both pass through the target location (2,2).

```{r TargetPathsPlot}
# Define the grid and paths
path1 <- rbind(c(0, 0), c(1, 0), c(2, 0), c(2, 1), c(2, 2), c(3, 2), c(4, 2), c(4, 3), c(4, 4))
path2 <- rbind(c(0, 0), c(0, 1), c(1, 1), c(1, 2), c(2, 2), c(2, 3), c(3, 3), c(3, 4), c(4, 4))

# Set up plot parameters
plot(NULL, xlim = c(0, 4), ylim = c(0, 4), xlab = "X", ylab = "Y", xaxs = "i", yaxs = "i", asp = 1)
abline(h = 0:4, v = 0:4, col = "lightgray", lty = 3)

# Plot the paths
lines(path1, col = "blue", lwd = 2)
lines(path2, col = "red", lwd = 2)

# Plot target location
points(2, 2, col = "green", pch = 19, cex = 1)
```


The total number of target paths to a fixed destination can be computed with the
tpaths function.

```{r}
tpaths(destination = c(5,4), target = c(3,2))
```


### Target Paths Probability Functions

The tpaths function and the mpaths function can be used to define probability
distribution function.

The dtpaths function is a probability mass function where the number of target 
paths passing through a target location is the event of interest while the sum of 
target paths across all coordinates on the grid is the sample space. The first
example below shows the symmetry on a square grid.

```{r dtpaths_symmetry}
dtpaths(c(5,5), c(2,3))
dtpaths(c(5,5), c(3,2))
```


The next example shows normalization for dtpaths

```{r dtpaths_normalization}
  probs_tpaths <- 0
  for (x in 0:10){
    for (y in 0:10){
      probs_tpaths <- probs_tpaths + dtpaths(c(10,10), c(x,y))
    }
  }
```


The ptpaths function is a cumulative distribution function for the dtpaths 
probability mass function.  The distribution is obtained by iterating through all 
possible target coordinates up to and including (ti, tj), computing dtpaths at 
each of these target coordinates, and summing the results.

```{r ptpaths_ex}
ptpaths(destination = c(8,8),target = c(4,4))
```


The qtpaths function is a quantile function which takes a probability p and a 
destination (i, j) and returns the coordinates for the target location (ti, tj) 
where the cumulative probability is equal to or just greater than p.  The wise
argument allows you to search for the first target location meeting the condition
in a "row"-wise manner or a "col"umn-wise manner.  The default is "col"

```{r qtpaths_ex}
qtpaths(destination = c(5,5), p =.5)
qtpaths(destination = c(5,5), p = .5, wise ="row")
```


The rtpaths function is a random variable function that generates random target 
coordinates from the distribution defined by the dtpaths probability mass function.

```{r rtpaths_ex}
rtpaths(destination = c(10,10), n = 3)
```

